---

title: "index"
output: html_document
date: "2024-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

# remove any warning messages from the code!!! - PROFESSIONALISM MARKS (warning = FALSE)

# make sure to include sanity checks - counting missing values etc 

library(tidyverse)
library(here)
library(ggplot2)
library(transformr)
library(gganimate)


```

## Project description
Friday 8th November 2024. Class demo live follow along coding. Sheffield weather data.

Data from this link: https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt

The URL for these pages is: 
https://izziecomber.github.io/Sheffield_weather_data/

The repo for these pages is:
https://github.com/izziecomber/Sheffield_weather_data.git


## Import the data 
```{r import}


loc = "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt"

# split values by "tabs" with \t (white space) and skip the first 7 rows (description of data)
# this approach doesn't work 

#df <- read.csv(data_loc, sep = "\t", skip = 7)

# will not import as text and clean

dat <- readLines(loc)

# change the word "provisional" to nothing, replace every instance

dat <- gsub("Provisional", "", dat)
dat <- gsub("\\*", "", dat)
dat <- gsub("\\#", "", dat)

# save new file

writeLines(dat, "cleaned.text")

# import data - lost the column names 

df <- read.table("cleaned.text", skip = 7, fill = TRUE)


```


## Wrangle the data 
```{r}

cols <- c("year", "month", "max_t", "min_t", "af", "rain", "sun")

colnames(df) <- cols

```

## visualise the data 
```{r}

# need max temp to be a number, use as.numeric

df$max_t <- as.numeric(df$max_t)

# ggplot, change transparency of lines (can show you the density)
# ensure you label the axis with a detailed description of the measure and include the unit of measurement
# include title and modify the range of the axis 

# canvas 
p <- ggplot(df, aes(x = month, y = max_t, group = year, colour = year))

# ink - added line showing this year's temperature data 
# add custom legend for the line of the current year (i.e., "2024 data shown in read")
p + geom_line(alpha = 0.1) + 
  geom_line(data = df %>% filter(year == 2024), colour = "red", linewidth = 2)



```
## Animated plot 

We are going to use gganimate to create an animated plot. Animation is nice but think carefully about whether it is distracting or not. Animation needs to be relevant/in service of the data. 

```{r animated}

# this section may not work because need to download package 

# go to gganimate and download 

library(gganimate)

# calculate average temp of each year
# pipe function (%>%) is only used for data frames 
df2 <- df %>%
  group_by(year) %>%
  summarise(avg_t = mean(max_t))

# create animated plot 
# need 3rd aes for animated plots (frame)
p2 <- ggplot(df2, aes(x = year, y = avg_t), frame = year)

# cumulative keeps it on the screen, "FALSE" would just have the points sort of jumping around the screen

anim <- p2 + 
  geom_point() + 
  transition_manual(year, cumulative = TRUE) +
  ggtitle("The year is: {frame+1882}")
# {frame+1882} adds 1882 to the frame number - allows you to add the correct year to the frame of the animation 
# run the show the animation 
anim 

# can change the speed, phase of the points etc - will change how impactful the plot is

# export as gif, then save the plot to be embedded into github pages 


```

